Bayesian with CLR

### 2021-07-19 scripts/48_16s_beta_diversity.Rmd
update

@@

dist_ <- vegdist(cts_fil, method = 'bray')
bc <- cmdscale(dist_, k = 2)

bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct() %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'grp') +
  labs(title = 'PCOA of healthy and CART patients') +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients.pdf')

bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct() %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'center') +
  labs(title = 'PCOA of healthy and CART patients center') +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients centetr.pdf')

@@

dist_ <- vegdist(cts_fil, method = 'bray')
bc <- cmdscale(dist_, k = 2)

@@

mp %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'pool') +
  labs(title = 'PCOA of CART patients') +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 5, height = 6)
```

>>>>

ttt <- all_pheno %>% 
  ungroup() %>% 
  filter(grp == 'healthy') %>% 
  distinct(sampleid)

@@

cts_fil <- cts %>% 
  filter(asv_key %in% keepa) %>% 
  select(sampleid, asv_key,count_relative ) %>% 
  spread(key = 'asv_key', value = 'count_relative', fill = 0) %>% 
  column_to_rownames('sampleid')
library(vegan)
dist_ <- vegdist(cts_fil, method = 'bray')
eigen <- pcoa(dist_)$values$Eigenvalues
percent_var <- signif(eigen/sum(eigen), 3)*100

bc <- cmdscale(dist_, k = 2)

bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct() %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'grp') +
  labs(title = 'PCOA of healthy and CART patients') +
  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients.pdf')

@@

dist_ <- vegdist(cts_fil, method = 'bray')
eigen <- pcoa(dist_)$values$Eigenvalues
percent_var <- signif(eigen/sum(eigen), 3)*100

bc <- cmdscale(dist_, k = 2)

@@

mp %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'pool', size = 3) +
  labs(title = 'PCOA of CART patients') +
  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 9, height = 9)
```

### 2021-07-19 scripts/54_rethinking_binary.Rmd
update

```{r}
# another model with the crs grade 0,1,>=2 
mcrs3 <- ulam(
    alist(
        crs3 ~ dordlogit( phi , cutpoints ),
        phi <- b*logdiv_s + a[location] ,
        b ~ dnorm( 0 , 1),  
        a[location] ~ dnorm( 0 , 0.5 ),
        cutpoints ~ dnorm( 0 , 1.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
plot(precis(mcrs3, depth = 2, prob = 0.95) ) 
# the cutpoints are not of much interet, this suggests that the diversity really isn't having a correlation?
```

@@

>>>>

```{r}
# another model with the crs grade 0,1,>=2 
# mcrs3 <- ulam(
#     alist(
#         crs3 ~ dordlogit( phi , cutpoints ),
#         phi <- b*logdiv_s + a[location] ,
#         b ~ dnorm( 0 , 1),  
#         a[location] ~ dnorm( 0 , 0.5 ),
#         cutpoints ~ dnorm( 0 , 1.5 )
#     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
# plot(precis(mcrs3, depth = 2, prob = 0.95) ) 
# the cutpoints are not of much interet, this suggests that the diversity really isn't having a correlation?
```

@@

    diff = abs( cols[[1]] -cols[[2]] )
    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
    return(ret)
  }) 
hist(prior_tox$p)
```

### 2021-07-19 scripts/57_16s_lefse.Rmd
update

annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts_$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) %>% 
  mutate(taxa_asv = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}|s__{species}|a__{asv_key}'))

@@

cmds <- tibble(
  fns = fns
) %>% 
  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1  -u 2 -o 1000000   ')) %>% 
  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 
  select(-fns) %>% 
  gather() %>% 
  select(value) %>% 
  write_csv('../data/amplicon/lefse/lefse_run_cmd_taxa.sh', col_names = F)

@@

>>>>

annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts_$asv_key) %>% 
  mutate(ordr =  if_else(is.na(ordr), str_glue('unknown_of_class_{class}'), ordr),
         family =  if_else(is.na(family), str_glue('unknown_of_order_{ordr}'), family),
         genus =  if_else(is.na(genus) , str_glue('unknown_of_family_{family}'), genus),
         species =  if_else(is.na(species) , str_glue('unknown_of_genus_{genus}'), species)) %>% 
  mutate(taxa_asv = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}|s__{species}|a__{asv_key}'))

@@

cmds <- tibble(
  fns = fns
) %>% 
  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1  -u 2 -o 1000000   ')) %>% 
  mutate(run_cmd = str_glue('run_lefse.py -l  4 {fns}.in  {fns}.res')) %>% 
  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 
  select(-fns) %>% 
  gather() %>% 
  select(value) %>% 
  write_csv('../data/amplicon/lefse/lefse_run_cmd_taxa.sh', col_names = F)

@@

```{r}
# run in terminal:
# bash /Users/daia1/projects/CART_microbiome/data/amplicon/lefse/lefse_run_cmd_taxa.sh
```

### 2021-07-19 scripts/63_CLR_transformation.Rmd
update

```{r}
library(compositions)
library(tidyverse)
```

```{r}
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  select(-center) %>% 
  spread(key = 'sampleid', value = 'relab', fill = 0) %>% 
  mutate(genus = if_else(is.na(genus), 'NA',genus)) %>% 
  column_to_rownames('genus')

clr_res <- clr(genus) %>% 
  as.data.frame()  %>% 
  rownames_to_column('genus') %>% 
  gather('sampleid','clr', names(.)[2]:names(.)[ncol(.)])

@@

>>>>

```{r}
library(compositions)
library(tidyverse)
library(ggpubr)
```
CLR works with counts, not relab. And also need to add a pseudocount

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')

library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('asv_annotation_blast_ag')
cts <- get_counts_subset(meta$sampleid)
cts_ <- cts %>% 
  select(asv_key, sampleid, count)

annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts_$asv_key)

cts_all <- cts_ %>% 
  left_join(annot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  group_by(sampleid, genus) %>% 
  summarise(cnt = sum(count)) %>% 
  spread('sampleid', 'cnt', fill = 0) %>% 
  filter(!is.na(genus)) %>% 
  column_to_rownames('genus')
```

```{r}
clr_res <- clr(cts_all + 0.5) %>% 
  as.data.frame()  %>% 
  rownames_to_column('genus') %>% 
  gather('sampleid','clr', names(.)[2]:names(.)[ncol(.)])

@@

```{r}
log10 <- read_csv('../data/amplicon/stool/combined_5_genera.csv') %>% 
  gather('genus', 'log10', Akkermansia:Ruminococcus)

two <- log10 %>% 
  inner_join(clr_5)

two %>% 
  arrange(log10) %>% 
  #filter(log10 > -5) %>% 
  ggscatter(x = 'log10', y = 'clr', facet.by = 'genus', alpha = 0.4,
             add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "pearson",  label.sep = "\n") )
```

### 2021-07-19 scripts/65_bayesian_with_CLR.Rmd
update

```{r}
gtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus  + a[location] ,
        ba ~ dnorm( 0 , 0.75),
        bb ~ dnorm( 0 , 0.75),
        be ~ dnorm( 0 , 0.75),
        bf ~ dnorm( 0 , 0.75),
        br ~ dnorm( 0 , 0.75),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 16)
precis(gtox, prob = 0.95)   
plot(precis(gtox, depth = 2, prob = 0.95) )  
# the  Ruminococcus looks more significant now
```

@@

```{r}
gcr <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus  + a[location] ,
        ba ~ dnorm( 0 , 0.25),
        bb ~ dnorm( 0 , 0.25),
        be ~ dnorm( 0 , 0.25),
        bf ~ dnorm( 0 , 0.25),
        br ~ dnorm( 0 , 0.25),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr, prob = 0.95)
plot(precis(gcr, depth = 2, prob = 0.95) )
```

>>>>

```{r}
gtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus  + a[location] ,
        ba ~ dnorm( 0 , 0.25),
        bb ~ dnorm( 0 , 0.25),
        be ~ dnorm( 0 , 0.25),
        bf ~ dnorm( 0 , 0.25),
        br ~ dnorm( 0 , 0.25),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 16)
precis(gtox, prob = 0.95)   
plot(precis(gtox, depth = 2, prob = 0.95) )  
# the  Ruminococcus looks more significant now
```

```{r}
# make this plot myself with a pre-defined range in the axis
df <- precis(gtox,  prob = 0.95)  %>% as.tibble()
df <- df %>%
  mutate(variable = c('Akkermansia', 'Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus'))
#remove row number 1 (The intercept)
df %>% 
  ggplot(aes(x = variable, y = mean)) +
  geom_point(shape = 15,
             size  = 4, width = 0.1,
             position = "dodge", color="black") +
  geom_errorbar(aes(ymin  = `2.5%`,
                    ymax  = `97.5%`),
                width = 0.1,
                size  = 0.7,
                position = "dodge", color="turquoise4") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Variables") + ylab("Coefficient with 95% CI") +
  coord_flip(ylim = c(-1.5, 1.5)) +
  geom_hline(yintercept = 0, color = "red", size = 1) +
   theme(axis.title = element_text(size = 0)) +
  theme(axis.text = element_text(size = 14)) +
  labs(title = 'Toxicity') +
  ggsave('../figs/amplicon/bayesian_genera_tox_CLR.pdf', width = 5, height = 4)
```

@@

```{r}
gcr <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus  + a[location] ,
        ba ~ dnorm( 0 , 0.25),
        bb ~ dnorm( 0 , 0.25),
        be ~ dnorm( 0 , 0.25),
        bf ~ dnorm( 0 , 0.25),
        br ~ dnorm( 0 , 0.25),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr, prob = 0.95)
plot(precis(gcr, depth = 2, prob = 0.95) )
```
```{r}
# make this plot myself with a pre-defined range in the axis
df <- precis(gcr,  prob = 0.95)  %>% as.tibble()
df <- df %>%
  mutate(variable = c('Akkermansia', 'Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus'))
#remove row number 1 (The intercept)
df %>% 
  ggplot(aes(x = variable, y = mean)) +
  geom_point(shape = 15,
             size  = 4, width = 0.1,
             position = "dodge", color="black") +
  geom_errorbar(aes(ymin  = `2.5%`,
                    ymax  = `97.5%`),
                width = 0.1,
                size  = 0.7,
                position = "dodge", color="turquoise4") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Variables") + ylab("Coefficient with 95% CI") +
  coord_flip(ylim = c(-1.5, 1.5)) +
  geom_hline(yintercept = 0, color = "red", size = 1) +
   theme(axis.title = element_text(size = 0)) +
  theme(axis.text = element_text(size = 14)) +
  labs(title = 'CR') +
  ggsave('../figs/amplicon/bayesian_genera_cr_CLR.pdf', width = 5, height = 4)

get_table_from_database('counts_ag')
```
